---
title: "Steps"
author: "Oliver Bobon"
date: "13 Jun 2015"
output: html_document
---

--
What is the mean total number of steps taken per day?
--
```{r, echo=TRUE}
activity <- read.csv("~/Desktop/activity.csv")
sum(activity$steps,na.rm=T)
```
--
What is the mean number of steps taken per day?
--
```{r, echo=TRUE}
mean(activity$steps, na.rm=TRUE)
```
--
Here are the total number of step and the mean number of steps shown in a histogram:
--
```{r, echo=TRUE}
dates = activity$date
agg = aggregate(steps ~ date, activity, sum)
dates = agg$date
steps = agg$steps
plot(agg,type="h",xlab="Date",ylab="Steps")
```
--
Here are the average steps taken across all days for each interval shown in a histogram:
--
```{r, echo=TRUE}
agg1 = aggregate(steps ~ interval, activity, mean)
plot(agg1,type="l",xlab="Interval",ylab="Steps")
```
--
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps [Interval and number of steps]?
--
```{r, echo=TRUE}
MaxSteps = data.frame(tapply(agg1[,2], agg1[,1], max))
MaxSteps[which.max(agg1[,2]),1]
```
--
Number of missing values:
--
```{r, echo=TRUE}
sum(is.na(activity$steps))
```
--
Missing Values
--
Now the na values are replaced with the mean number of steps for the specific interval.
A new dataset is created with those values replaced.
```{r, echo=TRUE}
library(plyr)
activityz = activity
aggz = agg1
test = join(activityz,aggz,by="interval")
for(n in 1:length(test$steps)) {
	if(is.na(test[n,1])) {test[n,1] = test[n,4]}
}
aggw = aggregate(steps ~ interval, test, mean)
```
--
With the replaced values, a new histogram is made:
--
```{r, echo=TRUE}
plot(aggw,type="l",xlab="Interval",ylab="Steps")
```
--
Report of average and median steps with and without missing values
--
```{r, echo=TRUE}
rep1 = aggregate(steps~date,activity,mean,na.rm=FALSE)
rep2 = aggregate(steps~date,activity,median,na.rm=FALSE)
rep3 = aggregate(steps~date,test,mean)
rep4 = aggregate(steps~date,test,median)
repa = cbind(rep1,rep2$steps)
repb = cbind(rep3,rep4$steps)
colnames(repa)=c("Date","average steps","median steps")
colnames(repb)=c("Date","average steps","median steps")
print(repa)
print(repb)
```

--
A panel plot containing a time series plot of the 5-minute interval and the avergage number of steps taken, averaged across all weekdays or weekend days.
--

```{r, echo=TRUE}
k = as.Date(agg$date[1])
test1 = cbind(test,wday="")
colnames(test1)=c("steps","date","interval","mean","wday")

wdayc = ""
for(n in 1:length(test1$steps)) {
  k = as.Date(test1$date[n])
  k = weekdays(k)
  if (k == "Saturday") {
    wdayc[n] = "Weekend"
  } else if (k=="Sunday") {
    wdayc[n] = "Weekend"
  }  else wdayc[n] = "Weekday" 
}
test2 = cbind(test1,wdayc)
subwd = subset(test2,wdayc=="Weekday")
subwe = subset(test2,wdayc=="Weekend")

aggwd = aggregate(steps ~ interval,subwd,mean)
aggwe = aggregate(steps ~ interval,subwe,mean)
par(mfrow=c(1,2))
plot(aggwd,type="l",xlab="Interval",ylab="Steps on Weekday")
plot(aggwe,type="l",xlab="Interval",ylab="Steps on Weekend")
```
